Cluster Difference Imaging Photometric Survey (CDIPS) II. TOI$\,$837: A Validated Jupiter-Sized Planet in IC~2602

### 2020-06-24 paper/ms.tex
minor manuscript updates

% TODO: SpType subtype?
\begin{abstract}
  We report the discovery of TOI 837.01 and its validation as a
  transiting planet.  We characterize the system using data from the
  NASA TESS mission, the ESA Gaia mission, ground-based photometry,
  and spectroscopy from CHIRON, FEROS, and Veloce.  We find that
  TOI$\,$837 is a $T=9.9$ G0 dwarf in IC 2602 (the ``Southern
  Pleiades'').  The star and validated planet are therefore 30$\pm$15
  million years old.  The planet is warm ($P = 8.32\,{\rm d}$) and
  roughly Jupiter-sized.  However, its transits are grazing: $b >
  0.92$ at 3$\sigma$.  From TESS photometry alone, the planetary size
  lies within 0.58--3.87 $R_{\rm Jup}$, due to the degeneracy between
  the planetary size and the impact parameter of the transit.  From
  radial velocity monitoring, we limit the mass of planet to less than
  2.XX $M_{\rm Jup}$ (3$\sigma$).  Grazing transits are cause for
  concern: they usually are indicative of astrophysical false positive
  scenarios.  Our follow-up data show that such scenarios are highly
  unlikely.  Multi-color photometry requires eclipsing companions to
  have nearly identical color as TOI$\,$837, ruling out hierarchical
  eclipsing binary scenarios.  Background eclipsing binary scenarios
  are limited by speckle imaging, but formally remain a 0.2\%
  possibility.  TOI 837.01 is therefore a validated adolescent
  exoplanet.  Further observations of its stellar obliquity,
  atmosphere, and mass should confirm its planetary nature beyond
  doubt, and could improve our understanding of how the physical and
  orbital properties of exoplanets change in time.
\end{abstract}

@@

Many of the exposures were taken in average or poor seeing conditions,
when fiber-to-fiber cross-contamination on the IFU-style fiber feed is
strongest.  To reduce the spectra to velocities, we cross-correlated
against a high S/N template of $\delta$ Pavonis.  The velocity RMS
seen across each visit was hundreds of meters per second, likely due to
uncorrected fiber-to-fiber cross-contamination.  For
analysis purposes, we averaged across each visit, and set the velocity
uncertainties as the standard deviation in the per-visit exposures.
The velocities are given in Table~\ref{tab:rvs}, and shown in
Figure~\ref{fig:rvs}.

@@

\subsubsection{RVs}
The combined radial velocities from FEROS, CHIRON, and Veloce can be
used to place limits on the presence of massive, close-in bounds
companion to \tn.  We derived the relevant limits using
\texttt{radvel} \citep{fulton_radvel_2018}.  We assumed circular
orbits for the massive companion.  We then fitted the RV data for the
semiamplitude, period, time of conjunction, instrument offsets, and
jitter parameters.  We placed a wide log-uniform prior on the
semiamplitude and period to ensure appropriate sampling of the
relevant spaces.  ($\log K \sim \mathcal{U}(1,10^5)$, $\log P \sim
\mathcal{U}(0.1, 10^{15})$).  We converted the resulting posterior in
period and semiamplitude to minimum mass and semi-major axis assuming
Kepler's third law.  The resulting $3\sigma$ limits are shown in
Figure~\ref{fig:fpscenario}.

@@

\subsubsection{Additional Transits}

None of the extra dips in the PDCSAP light-curve (e.g., ``up-down''
spikes at BTJD 1572 and 1601) seem likely to be planetary
We checked that i) they were not present in the SAPFLUX light-curves,
and ii) that they were not present in the CDIPS light-curves (either
raw, or PCA-detrended).

@@

\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  %\texttt{astroplan} \citep{astroplan2018},
	\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019}
  \texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{exoplanet} \citep{exoplanet:agol19}
  \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  dependencies \citep{exoplanet:agol19, exoplanet:kipping13, exoplanet:luger18,
  	exoplanet:theano}.
  \texttt{IPython} \citep{perez_2007},
	\texttt{isochrones} \citep{morton_2015_isochrones},
	\texttt{lightkurve} \citep{lightkurve_2018},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  %\texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
  \texttt{tesscut} \citep{brasseur_astrocut_2019},
	\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
  \texttt{webplotdigitzer} \citep{rohatgi_2019},
  \texttt{wotan} \citep{hippke_wotan_2019}.
}

>>>>

% TODO: SpType subtype?
\begin{abstract}
  We report the discovery of TOI 837.01 and its validation as a
  transiting planet.  We characterize the system using data from the
  NASA TESS mission, the ESA Gaia mission, ground-based photometry,
  and spectroscopy from CHIRON, FEROS, and Veloce.  We find that
  TOI$\,$837 is a $T=9.9$ G0 dwarf in IC 2602 (the ``Southern
  Pleiades'').  The star and validated planet are therefore 30$\pm$15
  million years old.  The planet is warm ($P = 8.32\,{\rm d}$) and
  roughly Jupiter-sized.  However, its transits are grazing: $b >
  0.92$ at 3$\sigma$.  From TESS photometry alone, the planetary size
  lies within 0.58--3.87 $R_{\rm Jup}$ (3$^{\rm rd}$--97$^{\rm th}$
  percentile), due to the degeneracy between the planetary size and
  the impact parameter of the transit.  From radial velocity
  monitoring, we limit the minimum mass to less than 1.20 $M_{\rm
  Jup}$ (3$\sigma$).  Grazing transits are cause for concern: they
  usually are indicative of astrophysical false positive scenarios.
  Our follow-up data show that such scenarios are highly unlikely.
  Multi-color photometry requires eclipsing companions to have nearly
  identical color as TOI$\,$837, ruling out hierarchical eclipsing
  binary scenarios.  Background eclipsing binary scenarios are limited
  by speckle imaging, but formally remain a 0.2\% possibility.  TOI
  837.01 is therefore a validated adolescent exoplanet.  Further
  observations of its stellar obliquity, atmosphere, and mass should
  confirm its planetary nature beyond doubt, and could improve our
  understanding of how the physical and orbital properties of
  exoplanets change in time.
\end{abstract}

@@

Many of the exposures were taken in average or poor seeing conditions,
when fiber-to-fiber cross-contamination on the IFU-style fiber feed is
strongest.  To reduce the spectra to velocities, we cross-correlated
against a high S/N template of $\delta$ Pavonis.  The velocity RMS
seen across each visit was hundreds of meters per second, likely due to
uncorrected fiber-to-fiber cross-contamination.  For
analysis purposes, we averaged across each visit, and set the velocity
uncertainties to be the standard deviation of the per-visit exposures.
The velocities are given in Table~\ref{tab:rvs}, and shown in
Figure~\ref{fig:rvs}.

@@

\subsubsection{RVs}
The combined radial velocities from FEROS, CHIRON, and Veloce can be
used to place limits on the presence of massive, close-in bound
companion to \tn.  We derived the relevant limits using
\texttt{radvel} \citep{fulton_radvel_2018}.  We assumed circular
orbits for the companion, and then performed two sets of fits.

% timmy.drivers.oneoff_drivers.get_mass_upper_limit
For the first fit, we assumed a restrictive prior on the period and
time of periapse, using the known ephemeris from the transit.  We then
fitted for the period, time of periapse, semi-amplitude, instrument
offsets, and jitter parameters.  This yielded a non-detection of the
planet's orbit.  The corresponding 3-$\sigma$ (99.7$^{\rm th}$
percentile) upper limit on $M_{\rm p} \sin i$ is $1.20 M_{\rm Jup}$.
The data and corresponding model are shown in Figure~\ref{fig:rvs}.

The above exercise rules out the possibility that the observed
eclipses are caused by a star orbiting \tn.  The lack of a linear
trend in the radial velocities, particularly in the FEROS data, can be
used to further limit the presence of a hierarchical binary system.
To constrain this second possibility, we placed a wide log-uniform
prior on the semi-amplitude and period ($\log K\ [{\rm m\,s^{-1}}]
\sim \mathcal{U}(1,10^5)$, $\log P\ [{\rm days}] \sim \mathcal{U}(0.1,
10^{15})$).  We then fitted again for the semi-amplitude, period, time
of conjunction, instrument offsets, and jitter parameters.  We
converted the resulting posterior in period and semi-amplitude to
minimum mass and semi-major axis assuming Kepler's third law.  The
resulting $3\sigma$ limits are shown in Figure~\ref{fig:fpscenario}.

@@

\subsubsection{Comments on Joint Transit and Radial Velocity Fitting}

The available data for \tn\ put us in an odd position. The transit
data provide a minimum size due to the grazing geometry; the radial
velocity data provide a maximum mass ($\approx$1.20 $M_{\rm jup}$).
We could combine the two with an empirically-calibrated planetary
mass-radius relation to force a more restrictive upper bound on the
planetary size \citep[{\it e.g.},][]{chen_probabilistic_2016}.  The
result of such an exercise would tautologically be that the planet
would be smaller than the largest known Jovian planets ($\approx
2R_{\rm Jup}$).  This would also imply that the impact parameter is
slightly below unity.  The planet's size, in our eyes, would remain
poorly constrained even if we were to do this exercise. {\bf So we
don't do it.}

\subsubsection{Additional Transits}
None of the extra dips in the PDCSAP light-curve (e.g., ``up-down''
spikes at BTJD 1572 and 1601) seem likely to be planetary
We checked that i) they were not present in the SAPFLUX light-curves,
and ii) that they were not present in the CDIPS light-curves (either
raw, or PCA-detrended).

@@

\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  %\texttt{astroplan} \citep{astroplan2018},
	\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019}
  \texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{exoplanet} \citep{exoplanet:agol19}
  \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  dependencies \citep{exoplanet:agol19, exoplanet:kipping13, exoplanet:luger18,
  	exoplanet:theano}.
  \texttt{IPython} \citep{perez_2007},
	\texttt{isochrones} \citep{morton_2015_isochrones},
	\texttt{lightkurve} \citep{lightkurve_2018},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  \texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
  \texttt{tesscut} \citep{brasseur_astrocut_2019},
	\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
  \texttt{webplotdigitzer} \citep{rohatgi_2019},
  \texttt{wotan} \citep{hippke_wotan_2019}.
}
